### 2019-03-22 1_introduction.tex
Edits.

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models, languages and runtime systems. These systems often provide
capabilities such as asynchronous execution of units of work, called
\emph{tasks}, coordinated in some manner with data movement and
synchronization. But while the features of these systems are widely
written about in the literature, a large and often difficult to answer
question is how different systems will perform for a given
application.

@@

Together, METG and Task Bench help to mitigate issues which are common
in limit studies on the task scheduling throughput of runtime
systems. Such studies typically employ a metric called \emph{tasks per
  second} (TPS), which is almost universally measured with trivial
dependencies and empty tasks. While TPS is certainly an upper bound on
task scheduling throughput, it is not a tight bound, as the cost of
non-trivial dependencies can be significant and sometimes adversely
impact scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

But there is a deeper issue with the use of empty tasks. In all
realistic scenarios, runtime systems must share resources with the
application, either explicitly by e.g., allocating cores for runtime
use, or implicitly by sharing a time slice on the same cores used for
application tasks, and other resources such as memory bandwidth and
processor cache capacity. The use of empty tasks makes the application
share of these resources appear to be zero, which permits usage by the
runtime system to be unconstrained. How then is one supposed to
determine what resource usage to consider reasonable on the part of
the runtime system? Only by requiring the application usage of
resources to meet a certain minimum level, as in METG, can we ensure
that runtime system resource usage is constrained to realistic levels.

>>>>

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models, languages and runtime systems. These systems often provide
capabilities such as the asynchronous execution of units of work, called
\emph{tasks}, coordinated in some manner with data movement and
synchronization. But while the features of these systems are widely
written about in the literature, a large and often difficult to answer
question is how different systems will perform for any given
application.

@@

Together, METG and Task Bench help to mitigate issues which are common
in limit studies on the task scheduling throughput of runtime
systems. Such studies typically employ a metric called \emph{tasks per
  second} (TPS), which is almost universally measured with trivial
dependencies. While TPS is certainly an upper bound on
task scheduling throughput, it is not a tight bound, as the cost of
non-trivial dependencies can be significant and sometimes adversely
impacts scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

But there is a deeper issue with the lack of an efficiency constraint
in measurements of TPS. TPS may be measured with empty
tasks~\cite{LegionTracing18}, or with tasks of some
duration~\cite{Canary16, Armstrong14}. If empty tasks are used to
measure TPS, this leads to an upper bound on the task scheduling
throughput, but fails to represent any realistic application, as no
useful work is being performed. If non-empty tasks are used, \emph{but
  the efficiency of the overall application is not reported}, then TPS
ceases to be a measurement of the limit of task scheduling
throughput. Large tasks may be used to hide any amount of overhead,
while tasks that are too small may result in a drop in total
application throughput even as TPS increases. Only by constraining
efficiency, as in METG, can we ensure a limit study that represents
reasonable throughput on useful work by the application.

### 2019-03-22 4_implementation.tex
Edits.

Internally, Chapel implements custom support for Cray networks, and
uses the GASNet~\cite{GASNET07} compatibility layer on other networks.

@@

Internally, Charm++ implements support for a variety of networks
directly on top of the low-level, vendor-provided APIs. For Cori, we
use Charm++'s native support for uGNI.

@@

In our evaluation, we use the default MPI implementation installed on
the machine, which is Cray MPI on the Cori supercomputer. This
implementation is vetted by the machine operators and vendor and is
the highest performing available implementation for the hardware.

@@

OmpSs~\cite{OmpSs11} is a programming model for loop-based parallelism
that is source-compatible with OpenMP. Our Task Bench implementation
uses OpenMP 4.0-style tasks with task dependencies. Because tasks in
OpenMP 4.0 have a fixed number of dependencies, we use a switch
statement to implement the dynamic dependencies required for Task
Bench. The implementation is otherwise straightforward.

@@

OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
parallelism on shared-memory systems. Our Task Bench implementation
for OpenMP is very similar to OmpSs and uses OpenMP 4.0 tasks with
task dependencies.

@@

@@

Internally, PaRSEC uses MPI for network communication.

@@

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model that is used internally by Legion~\cite{Legion12} and
Regent~\cite{Regent15}, which are implicitly parallel layers written
on top of Realm. Thus the Realm implementation of Task Bench can be
seen as a limit study of what is possible with Legion or Regent.

@@

Internally, Realm uses GASNet~\cite{GASNET07} for communication.

@@

Regent~\cite{Regent15} is an implicitly-parallel task-based language
which implements the Legion programming model~\cite{Legion12}. We use
Regent rather than Legion directly because the Regent compiler
provides a critical optimization, \emph{control replication}, that
enables superior horizontal scalability of
applications~\cite{ControlReplication17}.

@@

The Regent compiler performs control replication to slice the Regent
program into \emph{shards} of execution which can scalably execute on
large numbers of nodes. Our Regent implementation also makes use of a
tracing optimization~\cite{LegionTracing18} which substantially
reduces the cost of dynamic analysis of tasks in Regent.

@@

Spark's programming model is based of \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallelizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

Though Spark has tasks internally, these are not exposed to the user,
so one of the challenges in developing a Task Bench implementation is
mapping the task graph to a set of operators which will result in the
desired performance behavior. We use a combination of
\lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} operations to generate task
dependencies, and then \lstinline[language=Scala]{map} to execute a
series of tasks. We use an explicit hash partitioner to ensure that
Spark does not attempt to group multiple Task Bench tasks into a
single Spark task.

We performed an extensive set of experiments to verify that no
extraneous factors interfered with our Spark measurements. We
completely disabled logging, ensured that no RDDs are actually written
to disk, confirmed that there is no measurable overhead due to JNI
calls from Java to C, and cross-checked our results with known cases
that hit optimal task throughput in Spark, among other things. Our
results are discussed in detail in Section~\ref{sec:evaluation}.

@@

Internally, StarPU uses MPI for communication.

@@

Internally, Swift/T uses MPI for communication.

@@

Our Task Bench implementation uses
\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
between places, along with place-changing and atomic integers for
synchronization. We use the C++ backend of X10, which uses MPI
internally for portability to a variety of networks, and which
conveniently integrates well with the Task Bench API.

>>>>

@@

@@

@@

OmpSs~\cite{OmpSs11} is a programming model for loop- and task-based parallelism
that is source-compatible with OpenMP. Our Task Bench implementation
uses OpenMP 4.0-style tasks with task dependencies. Because tasks in
OpenMP 4.0 have a fixed number of dependencies, we use a switch
statement to implement the dynamic dependencies required for Task
Bench. The implementation is otherwise straightforward.

@@

OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
for OpenMP is very similar to OmpSs and uses tasks with
task dependencies.

@@

% FIXME: keep this in sync with experiments

@@

@@

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model that used internally by Legion~\cite{Legion12} and
Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
seen as a limit study of what can be achieved with Legion or Regent.

@@

@@

Regent~\cite{Regent15} is an implicitly-parallel task-based language
which implements the Legion programming model~\cite{Legion12}. We use
Regent rather than Legion directly because the Regent compiler
provides a critical optimization, \emph{control replication}, that
enables superior horizontal scalability~\cite{ControlReplication17}.

@@

The Regent compiler performs control replication to slice the Regent
program into \emph{shards} of execution which can scalably execute on
large numbers of nodes.

@@

Spark's programming model is based of \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallelizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that are cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

Though Spark has tasks internally, these are not exposed to the user,
so one of the challenges in developing a Task Bench implementation is
mapping the task graph to a set of operators which will result in the
desired performance behavior. We use a combination of
\lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} operations to generate task
dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
series of tasks. We use an explicit hash partitioner to ensure that
Spark does not attempt to group multiple Task Bench tasks into a
single Spark task.

We performed an extensive set of experiments to verify that no
extraneous factors interfered with our Spark measurements. We
completely disabled logging, ensured that no RDDs are actually written
to disk, confirmed that there is no measurable overhead due to JNI
calls from Java to C, and cross-checked our results with known cases
that hit optimal task throughput in Spark, among other things.

@@

@@

@@

Our Task Bench implementation uses
\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
between places, along with place-changing and atomic integers for
synchronization. We use the native backend of X10, which compiles to
C++.
